---
title: 'Stop and Frisk: Hypothesis Testing'
author: Howard Baik
date: '2019-08-30'
slug: stop-and-frisk-hypothesis-testing
categories:
  - R
tags:
  - Computational Social Science
subtitle: ''
summary: 'I will be performing an exploratory data analysis of the stop and frisk dataset provided by the New York Civil Liberties Union'
authors: []
lastmod: '2019-08-30T16:26:42+09:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

<br>

## Introduction

_From New York Civil Liberties Union:_

> Every time a police officer stops a person in NYC, the officer is supposed to fill out a form recording the details of the stop. The forms were filled out by hand and manually entered into an NYPD database until  2017, when the forms became electronic. The annual database includes nearly all of the data recorded by the police officer after a stop such as the age of the person stopped, if a person was frisked, if there was a weapon or firearm recovered, if physical force was used, and the exact location of the stop within the precinct.

<br>

In this post, I will be performing an exploratory data analysis of the stop-and-frisk dataset provided by the [New York Civil Liberties Union.](https://www.nyclu.org/en/stop-and-frisk-data) The data contains 11008 rows and 83 variables. Then, I will use hypothesis testing with the `infer` [package](https://infer.netlify.com/) to assess if there is a statistical difference in police action when approaching a black New Yorker vs a non-black New Yorker.

<br>

## Exploratory Data Analysis


```{r, include=FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
library(ggmap)
library(infer)
library(cowplot)
theme_set(theme_light())
register_google(key = "AIzaSyAqcHj2vykSo2ewu0LS1mdlHxNigBli1FY")

# Source of Data: https://www.nyclu.org/en/stop-and-frisk-data
sqf <- read_excel("2018_sqf_database.xlsx") %>% 
  janitor::clean_names() 

# Geocodes
geo_1_2500 <- read_csv("geocode_1_2500.csv")
geo_2501_5000 <- read_csv("geocode_2501_5000.csv")
geo_5001_7500 <- read_csv("geocode_5001_7500.csv")
geo_7501_10000 <- read_csv("geocode_7501_10000.csv")
geo_10001_end <- read_csv("geo_10001_end.csv")

geocodes <- geo_1_2500 %>% 
  rbind(geo_2501_5000) %>% 
  rbind(geo_5001_7500) %>% 
  rbind(geo_7501_10000) %>% 
  rbind(geo_10001_end)

# Get a map of manhattan
manhattan_map <- get_map(location = c(lon = -74.00, lat = 40.77), maptype = "terrain", zoom = 11)
```


```{r, echo=FALSE, message = FALSE}
sqf %>% 
  filter(stop_duration_minutes < 30) %>% 
  ggplot(aes(stop_duration_minutes)) +
  geom_histogram(fill = "cornflowerblue") +
  scale_x_continuous(breaks = c(0, 3, 5, 10, 15, 20, 25, 30)) +
  labs(x = "Stop Duration in Minutes",
       y = NULL) +
  theme(text = element_text(family = "Avenir")) +
  ggtitle("Stop duration tends to be in multiples of 5",
          subtitle = "Stop duration less than 30 minutes")
```

<br>


```{r, echo = FALSE, warning=FALSE}
p1 <- sqf %>% 
  count(suspected_crime_description, sort = TRUE) %>% 
  mutate(suspected_crime_description = if_else(suspected_crime_description =="CPW",
                                               "CRIMINAL POSSESSION OF WEAPONS",
                                               suspected_crime_description)) %>% 
  head(12) %>% 
  mutate(suspected_crime_description = fct_reorder(suspected_crime_description, n)) %>% 
  ggplot(aes(suspected_crime_description, n)) +
  geom_col(fill = "cornflowerblue") +
  theme(text = element_text(family = "Avenir")) +
  coord_flip() +
  labs(x = NULL,
       y = NULL) +
  ggtitle("Top 12 Suspected Crime Descriptions")


p2 <- sqf %>% 
  count(suspect_arrest_offense, sort = TRUE) %>% 
  filter(suspect_arrest_offense != "(null)") %>% 
  mutate(suspect_arrest_offense = if_else(suspect_arrest_offense =="CPW",
                                               "CRIMINAL POSSESSION OF WEAPONS",
                                               suspect_arrest_offense)) %>% 
  head(12) %>% 
  mutate(suspect_arrest_offense = fct_reorder(suspect_arrest_offense, n)) %>% 
  ggplot(aes(suspect_arrest_offense, n)) +
  geom_col(fill = "cornflowerblue") +
  theme(text = element_text(family = "Avenir")) +
  coord_flip() +
  labs(x = NULL,
       y = NULL) +
  ggtitle("Top 12 Suspect Arrest Offense")


plot_grid(p1, p2, ncol = 1, align = "hv")
```


```{r, echo = FALSE, message = FALSE}
sqf %>% 
  filter(suspect_reported_age != "(null)") %>% 
  mutate(suspect_reported_age = as.numeric(suspect_reported_age)) %>% 
  ggplot(aes(suspect_reported_age)) +
  geom_histogram(fill = "cornflowerblue") +
   geom_vline(xintercept = 26, color = "black", size = 2, lty = 2) +
  theme(text = element_text(family = "Avenir")) +
  labs(x = "Suspect Reported Age",
       y = NULL) +
  ggtitle("Suspect Reported Age is Skewed to the right",
          subtitle = "Median Age: 26") 
```


<br>


## Understanding Police Stops with Maps

```{r, echo=FALSE, warning = FALSE, message = FALSE}
geocodes_stop_duration <- geocodes %>%
  mutate(stop_duration_minutes = sqf$stop_duration_minutes) %>% 
  # Filter for points in Manhattan, Brooklyn, Bronx, Queens
  filter(between(lon, -74, -71),
         between(lat, 40, 41)) %>% 
  # Arrange so that redder colors appear on top of bluer colors
  arrange(stop_duration_minutes)



ggmap(manhattan_map) +
  geom_point(data = geocodes_stop_duration, aes(x = lon, y = lat, color = stop_duration_minutes)) +
  scale_color_gradient2(low = "blue",
                        high = "red",
                        midpoint = log10(median(sqf$stop_duration_minutes)),
                        trans = "log10",
                        labels = scales::comma_format()) +
  theme_void() +
  theme(text = element_text(family = "Avenir")) +
  labs(color = "Stop Duration (Min)",
       title = "Stop duration in New York City",
       subtitle = "Manhattan, Brooklyn, Bronx, Queens")
```

```{r, echo = FALSE, warning=FALSE, message=FALSE}
geocodes_stop_duration %>% 
  ggplot(aes(stop_duration_minutes)) +
  geom_histogram() +
  labs(x = "Stop Duration (Min)",
       y = NULL,
       title = "Skewed to the right")
```


- In this map, I'm using a couple tricks I learned from [David Robinson's screencasts](https://www.youtube.com/user/safe4democracy/videos). First, in order to better show the higher values of `stop_duration_minutes`, I arrange in increasing order of `stop_duration_minutes`. This allows ggplot to plot the lower values (lighter colors) first and then plot the higher values (redder colors) next, allowing me to emphasize regions with long stop duration in redder colors. Second, the histogram shows that the stop duration is skewed to the right- according to David, "you can go a little bit below the median, but way above". As a result, I transform the scale using the `trans` argument and choose the midpoint to be `log10(median(value))`. This way, the color scale gives more meaning to the data. 


```{r, echo=FALSE, warning=FALSE, message = FALSE}
geocodes_observed_duration <- geocodes %>%
  mutate(observed_duration_minutes = sqf$observed_duration_minutes) %>% 
  # Filter for points in Manhattan, Brooklyn, Bronx, Queens
  filter(between(lon, -74, -71),
         between(lat, 40, 41)) %>% 
  # Arrange so that redder colors appear on top of bluer colors
  arrange(observed_duration_minutes)



ggmap(manhattan_map) +
  geom_point(data = geocodes_observed_duration, aes(x = lon, y = lat, color = observed_duration_minutes)) +
  scale_color_gradient2(low = "blue",
                        high = "red",
                        midpoint = log10(median(sqf$observed_duration_minutes)),
                        trans = "log10",
                        labels = scales::comma_format()) +
  theme_void() +
  theme(text = element_text(family = "Avenir")) +
  labs(color = "Observed Duration (Min)",
       title = "Observed duration in New York City",
       subtitle = "Manhattan, Brooklyn, Bronx, Queens")
```


```{r, echo = FALSE, warning=FALSE, message=FALSE}
geocodes_observed_duration %>% 
  filter(observed_duration_minutes < 250) %>% 
  ggplot(aes(observed_duration_minutes)) +
  geom_histogram() +
  labs(x = "Observed Duration (Min)",
       y = NULL,
       title = "Skewed to the right")
```

<br>


```{r, echo=FALSE, warning=FALSE, message=FALSE}
geocodes_suspect_race <- geocodes %>%
  mutate(suspect_race_description = sqf$suspect_race_description) %>% 
  # Filter for points in Manhattan, Brooklyn, Bronx, Queens
  filter(between(lon, -74, -71),
         between(lat, 40, 41)) %>% 
  # Combine some races into one
  mutate(suspect_race_description = if_else(suspect_race_description == "BLACK HISPANIC", "BLACK",
                                            suspect_race_description),
         suspect_race_description = if_else(suspect_race_description == "WHITE HISPANIC", "WHITE",
                                            suspect_race_description)) %>%  
  filter(suspect_race_description != "(null)")


ggmap(manhattan_map) +
  geom_jitter(data = geocodes_suspect_race, aes(x = lon, y = lat, color = suspect_race_description),
              alpha = 0.25) +
  labs(color = "Suspect Race",
       title = "Suspect Races in New York City",
       subtitle = "Manhattan, Brooklyn, Bronx, Queens") +
  theme_void() +
  theme(text = element_text(family = "Avenir")) 
```

- We can see that suspects are predominantly white or black, with apparently more black suspects than white suspects. 


<br>

## Racial discrimination in frisking/searching suspects

I will perform hypothesis testing to find whether police engage in an action, specified by each of these columns, more on black suspects than non-black suspects in a statistically significant manner. Specfically, I will use the `infer` package, following the steps layed out in [Modern Dive Chapter 9](https://moderndive.com/9-hypothesis-testing.html)

```{r, echo=FALSE}
sqf %>% 
  filter(suspect_race_description != "(null)") %>% 
  mutate(suspect_race_description = if_else(suspect_race_description == "BLACK HISPANIC", "BLACK",
                                                         suspect_race_description)) %>% 
  mutate(suspect_race_description = if_else(suspect_race_description != "BLACK", "NON-BLACK",
                                                         suspect_race_description)) %>% 
  group_by(suspect_race_description) %>% 
  # Find proportion of Yes 
  summarise_at(vars(contains("physical"), contains("frisked"), contains("searched")), ~mean(. == "Y")) %>% 
  # Rename all columns except first column
  rename_at(-1, ~paste0(., "_prop")) %>% 
  gather("action", "proportion", -suspect_race_description) %>% 
  # Filter out non-existent actions
  filter(!(action %in% c("physical_force_weapon_impact_flag_prop",
                         "physical_force_oc_spray_used_flag_prop",
                         "physical_force_cew_flag_prop"))) %>% 
  mutate(action = case_when(
    action == "frisked_flag_prop" ~ "Frisked",
    action == "physical_force_draw_point_firearm_flag_prop" ~ "Draw Firearm",
    action == "physical_force_handcuff_suspect_flag_prop" ~ "Handcuff",
    action == "physical_force_other_flag_prop" ~ "Other Physical Force",
    action == "physical_force_restraint_used_flag_prop" ~ "Restraint Used",
    action == "physical_force_verbal_instruction_flag_prop" ~ "Verbal Instruction",
    action == "searched_flag_prop" ~ "Searched"
    
  )) %>% 
  ggplot(aes(action, proportion, fill = suspect_race_description)) +
  geom_col(position = "dodge") +
  labs(x = NULL,
       y = NULL,
       fill = "Suspect Race Description",
       title = "Distribution of police action on suspects",
       subtitle = "Blacks vs Non-Blacks") +
  facet_wrap(~action, scales = "free") +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_x_discrete(labels = NULL) +
  scale_fill_manual(values=c("#000000", "#f5f5dc"), 
                    labels = c("Black", "Non-Black")) +
  theme(text = element_text(family = "Avenir")) 
```


- This facetted plot shows that the police frisk, handcuff, use restraint and verbal instruction on black suspects more than non-black suspects

<br>

### Using `infer` for hypothesis testing

- Let's first create the processed version of `sqf` by taking out `(null)` inside `suspect_race_description` and lumping `suspect_race_description` into two groups of race: `BLACK` and `NON-BLACK`. 

```{r}
sqf_testing <- sqf %>% 
  filter(suspect_race_description != "(null)") %>% 
  mutate(suspect_race_description = if_else(suspect_race_description == "BLACK HISPANIC", "BLACK",
                                                         suspect_race_description)) %>% 
  mutate(suspect_race_description = if_else(suspect_race_description != "BLACK", "NON-BLACK",
                                                         suspect_race_description))
```

<br>

- Then, we use `specify` to formulate the response and explanatory variables. Let's use `frisked_flag` as the response variable and obviously, `suspect_race_description` as the explanatory variable. The argument `success = "Y"` reveals that we are interested in the proportion of "Y" in the `frisked_flag` column. 

```{r, eval = FALSE}
sqf_testing %>% 
  specify(formula = frisked_flag ~ suspect_race_description, success = "Y") 
```

<br>

- Now, we set meta data required for hyphotesis testing where we set `null = "independence"` for two sample (`BLACK` and `NON-BLACK`) hypothesis testing. 


```{r, eval = FALSE}
sqf_testing %>% 
  specify(formula = frisked_flag ~ suspect_race_description, success = "Y") %>% 
  hypothesize(null = "independence") 
```


<br>

- Here, we `generate` replicates of “shuffled” datasets assuming the null hypothesis is true. 

```{r, eval = FALSE}
sqf_testing %>% 
  specify(formula = frisked_flag ~ suspect_race_description, success = "Y") %>% 
  hypothesize(null = "independence") %>% 
  generate(reps = 1000, type = "permute")
```


<br>

- Now, we `calculate` the appropriate summary statistic for each of our 1000 shuffles, called the test statistic. According to Modern Dive, _"...since the unknown population parameter of interest is the difference in population proportions, the test statistic of interest here is the difference in sample proportions"_ We have 1000 values of `stat` and we assign this dataframe to `null_distribution`

```{r}
null_distribution <- sqf_testing %>% 
  specify(formula = frisked_flag ~ suspect_race_description, success = "Y") %>% 
  hypothesize(null = "independence") %>% 
  generate(reps = 1000, type = "permute") %>% 
  calculate(stat = "diff in props", order = c("BLACK", "NON-BLACK"))
```

<br>

- Here, we calculate the observed difference in proportion between blacks and non-blacks by using the same code as above except we remove `hypothesize` and `generate`

```{r}
obs_diff_prop <- sqf_testing %>% 
  specify(formula = frisked_flag ~ suspect_race_description, success = "Y") %>% 
  calculate(stat = "diff in props", order = c("BLACK", "NON-BLACK"))
```

<br>

- Now, we visualize the `null_distribution` (values of the difference in proportions assuming that there is no racial discrimination) with a histogram and then "add what happened in real-life" with a red line and shades. Here, the shaded region is the p-value:

> A p-value is the probability of obtaining a test statistic just as or more extreme than the observed test statistic assuming the null hypothesis is true.


```{r}
visualize(null_distribution, bins = 10) + 
  shade_p_value(obs_stat = obs_diff_prop, direction = "right")
```


<br>

- In the above graph, we clearly see that the p-value is extremely low. We calculate the exact p-value ("fraction that the null distribution is shaded") with `get_p_value`

```{r}
null_distribution %>% 
  get_p_value(obs_stat = obs_diff_prop, direction = "right")
```


<br>

- Below, we create a wrapper function that does all the steps we just covered so that we can iterate this hypothesis testing on all our desired columns.

```{r, eval = FALSE}
# Wrapper Function
run_hypothesis_test <- function(response_var) {
  
  f <- as.formula(
  paste(response_var, 
        "suspect_race_description", 
        sep = " ~ "))
  
  null_distribution <- sqf_testing  %>% 
    specify(formula = f, success = "Y") %>% 
    # "independence" for hypotheses involving two samples
    hypothesize(null = "independence") %>% 
    generate(reps = 1000, type = "permute") %>% 
    calculate(stat = "diff in props", order = c("BLACK", "NON-BLACK"))
  
  
  obs_diff_prop <- sqf_testing %>% 
    specify(formula = f, success = "Y") %>% 
    calculate(stat = "diff in props", order = c("BLACK", "NON-BLACK"))
  
  # Calculate P-value
  null_distribution %>% 
    get_p_value(obs_stat = obs_diff_prop, direction = "right") # Because P(BLACK) > P(NON-BLACK)
}


# Find desired variables
testing_names <- sqf %>% 
  select(contains("physical"), contains("frisked"), contains("searched")) %>% 
  names()

results <- testing_names %>% 
  map_df(~(data.frame(p_value = run_hypothesis_test(.x))),
         .id = "response_variable")
```


```{r, include = FALSE, warning=FALSE, message = FALSE}
results <- read_csv("results.csv")
```


```{r, echo = FALSE}
results %>% 
  mutate(response_variable = fct_reorder(response_variable, p_value)) %>% 
  ggplot(aes(response_variable, p_value)) +
  geom_point() +
  geom_hline(yintercept = 0.05, size = 2, lty = 2) +
  coord_flip() +
  labs(x = NULL,
       y = "P-value",
       title = "Which variables have a p-value greater than alpha?",
       subtitle = "Alpha = 0.05")
```

> All response variables have p-values that are bigger than 0.05 except for `physical_force_restraint_used_flag` and `frisked_flag`. This signifies that we reject the null hypothesis for these variables and conclude that there is evidence to suggest that police frisk and use physical force on blacks suspects more than non-black suspects.

